Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2016-01-19 latex/article.tex
minor edits

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.
\color{blue}
Here, we focus on curvature tensor estimators.
Most of them require a parameter fixing the scale
at which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies a scale for analyzing the object surface, 
and is naturally related to the amount of perturbation on input data. As a consequence, when using
such tensors, exploring different values of this parameter is mandatory.
However, this is usually an offline process, due to the amount of computations that
needs to be done.
\normalcolor
%In this work, we present an interactive visualization of volumetric data isosurfaces combined with
%a curvature tensor (mean/Gaussian curvature, principal curvatures
%directions, \ldots) estimation at each point of the digital
%surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.

@@

\vspace{0.2cm}
\color{blue}
In this paper, we will first present the previous works on
curvature tensor estimation. Then, we will see how we can
efficiently extract a triangulation of an object on the GPU
and finally, we will present how we can evaluate the curvature tensor in
real time and discuss on the several approaches that we tried.
\normalcolor

@@

In \cite{DGCI2013}, estimators are defined based on the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $\VolC(Y,h) \EqDef h^d \MCard(Y)$ (we denote such
quantity $\AreaC(Y,h)$ in dimension 2):
%% Redite
%% \Jaco{Given some point $\vx \in \R^d$, the set $Y$ is the digital intersection of $\DigShape$ with the ball of radius $R/h$ and center $\vx/h$.}
%% where $Y$ is the intersection of $\DigShape$ with a ball
%% $\Ball{R/h}{\vx/h}$ with $\MCard(Y)$, the number of digital points of
%% the intersection between a ball $\Ball{R/h}{\vx/h}$ of digital radius
%% $R/h$ and $\DigShape$ (we denote such quantity $\AreaC(Y,h)$ in
%% dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
Estimators of both 2D curvature and 3D mean curvature on digital
objects with volumetric integral information are proved to be multigrid convergent
property \cite{DGCI2013}. For short, when the digital object becomes finer and
finer, \ie when the digitization step $h$ tends to zero, the estimated quantity
on $\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to the
\color{blue}
associated one on $\Bd{\Shape}$ in $O\left(h^{\frac{1}{3}}\right)$ for
convex shapes with at least $C^3$-boundary and bounded curvature.
\normalcolor
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
%Setting $R = kh^\frac{1}{3}$ on convex shapes
%with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
%related to the maximal curvature of the shape, they have:
%%
%\begin{align}
%\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
%&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
%\end{align}

@@

\color{blue}
To implement the integration on $\Ball{R}{\vx} \cap \Shape$ (Fig.~\ref{fig:approx}-$a$), we tried several approaches.

@@

\paragraph{Naive Approach.}
A first simple solution , consists in scanning all digital
\normalcolor
points inside the integration domain (Fig.~\ref{fig:approx}-$b$) and then to estimate the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection. 
\color{blue}
This is exactly similar to what is done on the CPU, but on the GPU, we can
exploit \emph{mipmap} textures to obtain a multi-resolution information.
\normalcolor
If the input binary object is stored in a
3D texture, GPU hardware constructs a multi-resolution pyramid (mipmap) such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
\color{blue}
a given level $l$, a texture probe returns a voxel $(x,y,z)$, containing the fraction of $\DSh$ belonging
\normalcolor
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$c,d$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a kind of adaptive Gauss digitization process.

@@

 \begin{figure}[h!]
   \vspace{-0.35cm}
   \begin{center}
   	 {\includegraphics[width=0.4\textwidth]{figs/snow_r8}}
   	 {\includegraphics[width=0.4\textwidth]{figs/xyzrgb_dragon-510_R8_mean}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_min}}
	 {\includegraphics[width=0.4\textwidth]{figs/dragon_r8}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_max}}
	 {\includegraphics[width=0.4\textwidth]{figs/dragon_r8_max}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_normales}}
     {\includegraphics[width=0.4\textwidth]{figs/dragon_r8_normales}}

>>>>

Volumetric objects are being more and more popular in many
applications ranging from object modeling and rendering in Computer
Graphics to geometry processing in Medical Imaging or Material
Sciences. When considering large volumetric data, interactive
%visualization of such images (or isovalue surfaces) is a complex
visualization of those objects (or isosurfaces) is a complex
problem. Such issues become even more difficult when dynamic
%volumetric datasets are considered. Beside such isosurface
volumetric datasets are considered. Beside
visualization, we are also interested in performing geometry
processing on the digital object and to explore different parameter
settings of the geometry processing tool.
\color{blue}
Here, we focus on curvature tensor estimation (mean/Gaussian curvature, principal curvatures
directions, \ldots). Most curvature estimators require a parameter fixing the scale
at which the computation is performed. For short, such parameter
(integration radius, convolution kernel size, \ldots) specifies a scale for analyzing the object surface, 
and is naturally related to the amount of perturbation on input
data\todo{DC: ``related to a expected signal/noise ratio?''}. As a consequence, when using
such estimators, exploring different values of this parameter is mandatory.
However, this is usually an offline process, due to the amount of computations that
needs\todo{DC: needs->need?} to be done.
\normalcolor
%In this work, we present an interactive visualization of volumetric data isosurfaces combined with
%a curvature tensor (mean/Gaussian curvature, principal curvatures
%directions, \ldots) estimation at each point of the digital
%surface.
%Curvature maps on a given isosurface are key information when
%exploring 3D structures in Material Sciences \todo{REF}.

@@

\vspace{0.2cm} \color{blue} In this paper, we first present the
previous works on curvature tensor estimation. Then, we propose na
efficient GPU approach to extract a triangulated isosurface from a
digital object. Finally, we propose a fully-parallel  GPU pipeline to compute
curvature tensor in real time and discuss on experiments. \todo{DC:
  bof le discuss}
\normalcolor

@@

In \cite{DGCI2013}, estimators are defined based on the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $\VolC(Y,h) \EqDef h^d \MCard(Y)$ (we denote such
quantity $\AreaC(Y,h)$ in dimension 2):
%% Redite
%% \Jaco{Given some point $\vx \in \R^d$, the set $Y$ is the digital intersection of $\DigShape$ with the ball of radius $R/h$ and center $\vx/h$.}
%% where $Y$ is the intersection of $\DigShape$ with a ball
%% $\Ball{R/h}{\vx/h}$ with $\MCard(Y)$, the number of digital points of
%% the intersection between a ball $\Ball{R/h}{\vx/h}$ of digital radius
%% $R/h$ and $\DigShape$ (we denote such quantity $\AreaC(Y,h)$ in
%% dimension 2):
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature
  estimators} are defined for any point $\vx \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\Ball{R/h}{\vx/h}$ is the ball of digital radius $R/h$ centered on
  digital point $(\vx/h) \in Z$.
  %
  % \AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  % \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  % \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  % the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  % the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  % $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
Estimators of both 2D curvature and 3D mean curvature on digital
objects with volumetric integral information are proved to be multigrid convergent
property \cite{DGCI2013}. For short, when the digital object becomes finer and
finer, \ie when the digitization step $h$ tends to zero, the estimated quantity
on $\Bd{\Body{\DSh}{h}}$ converges (theoretically and experimentally) to the
\color{blue}
associated one on $\Bd{\Shape}$ in $O\left(h^{\frac{1}{3}}\right)$ for
convex shapes with at least $C^3$-boundary and bounded curvature
(setting $R\EqDef kh^\frac{1}{3}$ for some $k\in\R$).
\normalcolor
%
%Since our digital estimators are defined on
%digital shapes, in order to compare results with the exact curvature on
%Euclidean shape, we need to match points on boundary of the digital
%shape and points on the border of the Euclidean one.
%
%Setting $R = kh^\frac{1}{3}$ on convex shapes
%with at least $C^3$-boundary and bounded curvature, where $k$ is a constant
%related to the maximal curvature of the shape, they have:
%%
%\begin{align}
%\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
%\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
%&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
%&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
%\end{align}

@@

\color{blue}
To implement the integration on $\Ball{R}{\vx} \cap \Shape$ (Fig.~\ref{fig:approx}-$a$), we tried several approaches.
\normalcolor

@@

\paragraph{Naive Approach.}
\txtblue{A first simple solution  consists in scanning all digital
points, at a given resolution, lying inside the integration domain}
(Fig.~\ref{fig:approx}-$b$) and then \txtblue{estimating} the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection (see
Eq.~(\ref{eq:dig-curvature-estimator-k}) to (\ref{eq:dig-curvature-estimator-k2})). 
\color{blue}
This is exactly similar to what is done on the CPU. On  the GPU, we can
exploit \emph{mipmap} textures to obtain a multi-resolution information.
\normalcolor
If the input binary object is stored in a
3D texture, GPU hardware constructs a multi-resolution pyramid (mipmap) such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
\color{blue}
a given level $l$, a texture probe returns a voxel $(x,y,z)$, containing the fraction of $\DSh$ belonging
\normalcolor
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$c,d$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a kind of adaptive Gauss digitization process.

@@

 \begin{figure}[h!]
   \vspace{-0.35cm}
   \begin{center}
   	 {\includegraphics[width=0.45\textwidth]{figs/snow_r8}}
   	 {\includegraphics[width=0.45\textwidth]{figs/xyzrgb_dragon-510_R8_mean}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_min}}
	 {\includegraphics[width=0.45\textwidth]{figs/dragon_r8}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_max}}
	 {\includegraphics[width=0.45\textwidth]{figs/dragon_r8_max}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_normales}}
     {\includegraphics[width=0.45\textwidth]{figs/dragon_r8_normales}}
